A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2021-11-15 analysis/paper/paper.Rmd
Struggeling with some strange crashes

For the date ranges associated with the sites, all radiocarbon dates were first individually calibrated using the IntCal20 calibration curve using OxCal v. . All dates not younger than 1700 BCE with 99.7 % uncertainty range were then excluded from further analysis. Radiocarbon dates associated with each site were then inspected, and with reference to the evaluation done in the original reports, dates believed to belong to the same settlement phase were grouped. In the case were there are multiple dates believed to belong to a single phase, these were subjected to Bayesian modelling using the Boundary function in OxCal. Multiple phases at a single site were treated as independent of each other (Figure x).

@@

Armed with a likely date range for the occupation(s) of each site, the likely trajectory of relative sea-level change at that location, and an DTM edited to remove substantial modern disturbances, the simulation was performed. Each simulation run involved first drawing a single year from the posterior density estimate of a given occupation phase. This year then has a corresponding likely elevation range of the shoreline. An elevation value was drawn uniformly from this range, using intervals of 5 cm. The sea-level was then raised to this elevation on the DTM by defining all elevation values at or below this altitude as missing. Polygons were then created from the resulting areas with missing values. The horisontal distance was then found by measuring the shortest distance between site and sea polygons, and the vertical distance by subtracting the elevation of the sea-level from the lowest elevation of the site polygon. The so-called topographic distance between site and sea was also found by measuring the distance while taking into account the slope of the terrain on the DTM. This was done using the topoDist package for R. In the case where the sea-polygons intersects the site polygon, all distance measures were set to zero. In the case that the sea-polygons completely contain the site, the horisontal and topographic distance measures were made negative. While its safe to assume that an archaeological site was not occupied when it was located beneath sea-level, a negative result can reflect the inherent uncertainty in this procedure and might also help identify discrepancies in displacement data or radiocarbon dates.

>>>>

For the date ranges associated with the sites, all radiocarbon dates were first individually calibrated using the IntCal20 calibration curve using OxCal v. . All dates not younger than 1700 BCE with 99.7 % uncertainty were then excluded from further analysis. Radiocarbon dates associated with each site were then grouped if they overlaped with 99.7 % uncertainty, effectively taken to represent the same settlement phase. In the case were there are multiple dates believed to belong to a single phase, these were subjected to Bayesian modelling using the Boundary function in OxCal. Multiple phases at a single site were treated as independent of each other (Figure x).

@@

Armed with a likely date range for the occupation(s) of each site, the likely trajectory of relative sea-level change at that location, and an DTM edited to remove substantial modern disturbances, the simulation was performed. Each simulation run involved first drawing a single year from the posterior density estimate of a given occupation phase. This year then has a corresponding likely elevation range of the contemporaneous shoreline. An elevation value was drawn uniformly from this range, using intervals of 5 cm. The sea-level was then raised to this elevation on the DTM by defining all elevation values at or below this altitude as missing. Polygons were then created from the resulting areas with missing values. The horisontal distance was then found by measuring the shortest distance between site and sea polygons, and the vertical distance by subtracting the elevation of the sea-level from the lowest elevation of the site polygon. The so-called topographic distance between site and sea was also found by measuring the distance while taking into account the slope of the terrain on the DTM. This was done using the topoDist package for R. In the case where the sea-polygons intersects the site polygon, all distance measures were set to zero. In the case that the sea-polygons completely contain the site, the horisontal and topographic distance measures were made negative. While its safe to assume that an archaeological site was not occupied when it was located beneath sea-level, a negative result can reflect the inherent uncertainty in this procedure and might also help identify discrepancies in displacement data or radiocarbon dates.
